### 2014-11-07 documents/paper/ms_v0.0.tex
current doc

\documentclass[12pt,preprint]{aastex}
\usepackage{amssymb,amsmath,mathrsfs,hyperref,datetime}

\newcommand{\datavector}[1]{\boldsymbol{#1}}
\newcommand{\data}{\datavector{y}}
\newcommand{\datum}{\data_i}
\newcommand{\truedatum}{\data_{{\rm true}, i}}
\newcommand{\noiselessdata}{\tilde{\data}}
\newcommand{\noiselessdatum}{\tilde{\datum}}
\newcommand{\xdcov}{\datavector{V}_j}
\newcommand{\datacov}{\datavector{C}}
\newcommand{\postcov}{\tilde{\datacov}_{i, j}}
\newcommand{\vmu}{\datavector{\mu}}
\newcommand{\postmu}{\tilde{\vmu}_{i, j}}
\newcommand{\datumcov}{\datacov_i}
\newcommand{\transpose}{{\rm T}}
\newcommand\rf[1]{{\bf [RF: #1]}}

\def\urltilda{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}

\pdfoutput=1

\begin{document}

\title{Joint photometric modeling for improved measurements in imaging surveys.}
\author{Ross~Fadely\altaffilmark{1} \&
        David~W.~Hogg\altaffilmark{1,2}}
\altaffiltext{1}{Center~for~Cosmology~and~Particle~Physics, Department~of~Physics, New~York~University, 4~Washington~Place, New~York, NY 10003, USA}
\altaffiltext{2}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, 69117 Heidelberg, Germany}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Abstract
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Large scale astronomical imaging surveys have fundamentally changed our 
understanding of the cosmos, and promise to continue to do so with forthcoming 
surveys like the Large Synoptic Survey Telescope which will provide greater 
volumes of data at increasingly fainter detection limits.  In order to extract
the most science out of such costly missions, it is important to continually
explore ways to improve the means by which photometric measurements are made.
Historically, photometric measurements are made on a per object basis under
our understanding of the characteristics of the telescope, instruments, and
astronomical sources.  We argue that it is possible to improve the quality of
such photometric measurements by building models which incorporate the shared
similarity of sources.  As a demonstration, we use Extreme Deconvolution (XD)
to model the joint distribution of photometric quantities measured by SDSS in
Stripe 82.  Using single epoch data, we show that our joint XD model makes
predictions that more closely resemble the coadded Stripe 82 data.  Looking
at the difference between the psf and model magnitudes in the single epoch and 
coadded data, the average error is XX at $r\sim21$ versus an average of XX
for our XD posteriors.  In the case of the CMDs of globular clusters, we show
that the XD posteriors lessen the presence of misclassified galaxies and
present a narrower stellar locus.  We suggest a number of improvements that
can be made to provide improved photometric inferences.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section - Introduction
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Large imaging surveys of parts (or the full) sky are one main modes in which
modern astronomers study the universe.  Whether conducted on the ground (e.g.,
SDSS, UKIDSS, NEWFIRM, VLA, etc.) or in space (GALEX, WISE, XMM?, etc.), the
volume of such surveys provides both strong statistical samples of objects and
also the potential for new discoveries.  For instance, the Sload Digital Sky
Survey (optical $ugriz$, $r\lesssim 22$) \citep{york00} was first dataset able
to describe the quasar luminosity function (CITE?) while also enabling the 
discovery of new streams and satellites around the Milky Way (CITE).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section - Method
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Extreme Deconvolution}

\begin{itemize}
\item Mixture of Gaussians where each datum's observational uncertainty is accounted for.
\item cite bovy and applicable works.
\item regularization to avoid singularities
\end{itemize}

\subsection{Posterior estimation}

After inference, the XD mixture provides an estimation for the noiseless distribution of 
the data:

\begin{eqnarray}\displaystyle
p(\noiselessdata) = \sum_j^K \frac{\alpha_j}{\sqrt{2\pi\det{|\xdcov|}}} \exp \left (-\frac{1}{2}(\noiselessdata - \vmu_j)^\transpose \xdcov^{-1}(\noiselessdata - \vmu_j) \right )
\quad .
\label{eqn:noiseless}
\end{eqnarray}

In a hierarchical sense, this learned distribution provides a prior distribution on what the 
value of a datum should be --- the distribution of the denoised data $\noiselessdata$ is 
learned for all the observations via XD.  Using Bayes' Rule, we can estimate the posterior 
value for a given datum

\begin{eqnarray}\displaystyle
p(\truedatum | \datum) & \propto & p(\datum | \truedatum) p(\truedatum) \\
                                         & \propto & \frac{1}{\sqrt{2\pi\det{|\datumcov|}}} \exp \left (-\frac{1}{2}(\datum - \truedatum)^\transpose \datumcov^{-1}(\datum - \truedatum) \right ) p(\truedatum)
\quad ,
\label{eqn:posterior}
\end{eqnarray}

\noindent where we assume that the prior for $\truedatum$ is $p(\noiselessdata)$.  Since 
$p(\noiselessdata)$ is a sum of Gaussian densities, and we multiply this by a Gaussian 
likelihood, the posterior is also a sum of Gaussians.  That is, the posterior for the true 
value of the observation is

\begin{eqnarray}\displaystyle
p(\truedatum | \datum) & \propto & \sum_j^K \frac{\alpha_j}{\sqrt{2\pi\det{|\postcov|}}} \exp \left (-\frac{1}{2}(\truedatum - \postmu)^\transpose \postcov^{-1}(\truedatum - \postmu) \right ) \\
{\rm where,}\\
\postcov & = & (\datumcov^{-1} + \xdcov^{-1})^{-1} \quad {\rm and,} \\
\postmu & = &  \postcov^{-1} (\datumcov^{-1} \datum + \xdcov^{-1} \vmu_j)
\label{eqn:posterior}
\end{eqnarray}
\end{document}